---
title: "Get started with {nmhsa}"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nmhsa)
```

## The example microstructure

The {nmhsa} package comes with an example microstructure, *i.e.* a segmented
cement paste image, with labels 1:4 corresponding to pores and mineral phases
portlandite, clinker and CSH respectively:

```{r}
cement
plot(cement)
```

This is just a conventional R matrix / 2D array, with an S3 class
("`nmhsa_array`") for more useful printing and quick plotting (well, for images
with up to nine phases). Use the `as.nmhsa_array()` function for adding the
class to any other matrix or array. Just try to make sure the values start
with 1.

## An overview of the algorithms

We provide the three 2D examples of the original Python code here, sometimes
maybe relying on slightly different defaults. All these can be easily extended
to 3D reconstruction by providing the dimensions for the reconstruction
explicitly, instead of relying on the default behaviour, which generates images
of the same size as the training image.

### Hierarchical simulated annealing (HSA)

We only set `order` explicitly here, and rely on the default parameters
otherwise, as it does not exactly correspond to the default `order`, which is
from the least to the most occurring phase:

```{r}
#| cache: true
reconstruction <- hsa(cement, order = c(3, 1, 2, 4))
plot(cement, reconstruction)
```

### Multiresolution hierarchical simulated annealing (MHSA)

To extend this to the multiresolution approach, we additionally provide the
`levels`:

```{r}
#| cache: true
reconstruction <- mhsa(cement, order = c(3, 1, 2, 4), levels = c(3, 2, 2, 1))
plot(cement, reconstruction)
```

### Nested multiresolution hierarchical simulated annealing (NMHSA)

The NMHSA example is a bit more complex, as we need to preprocess the training
image to provide information on the phase merging and subphase splitting. We
first start by merging neighbouring pixels of phase 3 into phase 2:

```{r}
cement_merged <- cement |>
  phase_merge(3, into = 2)
plot(cement, cement_merged)
```

Then we continue processing the image, and split off a series of subphases:

```{r}
cement_splitted <- cement_merged |>
  phase_split(1, larger_than = 3) |> # large pores
  phase_split(2, larger_than = 3) |> # large portlandite components
  phase_split(3, larger_than = 3) |> # large clinker components
  phase_split(7, larger_than = 15)   # extra large clinker components
plot(cement, cement_splitted)
```

Having these two modified versions of the training image, we can start the
reconstruction:

```{r}
#| cache: true
reconstruction <- nmhsa(
  ti = cement,
  ti_merged = cement_merged,
  ti_splitted = cement_splitted,
  order = c(8, 7, 5, 6, 3, 1, 2, 4),
  levels = c(3, 2, 2, 2, 1, 1, 1, 1),
  cool = 0.98,
  merge_pairs = list(c(3, 4), c(1, 4), c(2, 4))
)
plot(cement, reconstruction)
```
